\hspace{1cm} Do syntactic and semantic similarity lead to interference effects?\newline Evidence from self-paced reading and event-related potentials using German

### 2023-10-08 paper/main.tex
minor edits

\paragraph{Bayes factors for model comparison}
Model comparisons were carried out using Bayes factors. Because Bayes factors can be very sensitive to prior specifications on the target parameter being tested \parencite{schad_etal_2022_BF}, 
we report a sensitivity analysis using a range of prior specifications for the relevant parameters (see Table \ref{tab:spr_priors}). The priors assume a priori effect sizes for the main effects of syntactic and semantic interference, and their interaction, ranging from 0-8 ms, 0-40 ms, or 0-81 ms.  These priors are based on the observed effect sizes and their uncertainties in previous reading studies that use the present design \parencite{vandyke07,mertzen}.

@@

Bayes factors can be interpreted as follows \parencite[e.g., ][]{lee2014bayesian}: if BF$_{10}$ > 1, it provides evidence in favor of the effect of interest. If BF$_{10}$ < 1, it provides evidence in favor of the null model. The larger the value of BF$_{10}$, the stronger the evidence for the effect and the smaller the value of BF$_{10}$ is, the stronger the evidence for the null. 
In general, a large number of iterations is needed in the brms package in order to obtain stable estimates of the Bayes factor  \citep{schad_etal_2022_BF}.
For this reason, models were run with 4 chains and 20,000 iterations in each chain, with the first 2,000 iterations in each chain being discarded as the warm-up phase.  Bayes factors were unstable when varying slopes were included; for this reason, we fit the models only with varying intercepts by participant and by item.

@@

 %\multirow{3}{1cm}{lb=0}

@@

\subsubsection{Comprehension question accuracy}
After exclusion of participants with accuracy below 70\,\%, overall accuracy (including fillers) was good; on average 85.9\,\% (range: 71.9 - 100\,\%). Accuracy in critical trials was 81.1\,\% (range: 20 - 100\,\%). Condition-wise accuracy is shown in Table \ref{tab:spr_acc}.

\begin{table}[]
    \caption{By-condition accuracy in critical trials in the SPR experiment.}
    \label{tab:spr_acc}
    \centering
    \begin{tabular}{llr}
    \toprule
    syntactic & semantic & accuracy \%\\
    \midrule
        low &  low & 86.0\\
        low &  high & 77.7\\
        high &  low & 85.8\\
        high &  high & 74.9\\
    \bottomrule
    \end{tabular}
\end{table}

Table \ref{tab:spr_acc_mod} shows the results of the generalized mixed model analyzing the comprehension accuracy in critical trials. The estimates and 95\,\% credible intervals in log-odds show primarily a reduction in comprehension accuracy for high compared to low semantic interference conditions.

\begin{table}[H]
    \caption{Results in log-odds from the Bayesian generalized model analyzing the comprehension accuracy in critical trials of the SPR experiment.}
    \label{tab:spr_acc_mod}
    \centering
    \begin{tabular}{lrrr}
    \toprule
    & Estimate & Est.Error & 95\% CrI  \\
    \midrule
Intercept& 1.66 & 0.14&  1.38 - 1.93\\
syntactic& -0.08 & 0.04& -0.16 - -0.01\\
semantic&  -0.61 &0.04& -0.69 - -0.54\\
interaction& -0.10& 0.06& -0.22 - 0.02\\
    \bottomrule
    \end{tabular}
\end{table}

\subsubsection{Self-paced reading times}
Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence} A and B, separately for high and low syntactic interference because these conditions had partially different sentence structure. 
%\clearpage

@@

It is apparent from Figure \ref{fig:whole_sentence} A and B that, at the distractor, regardless of the syntactic manipulation, reading times between high and low semantic interference started to differ. Distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions. This is probably due to encoding interference \citep{mertzen, jaeger_etal_2015_retrieval_encoding} and persisted in the following regions. Nonetheless, we refrained from analyzing the distractor and the immediately following regions, because they differed between conditions. Instead we focused our analyses on regions which were identical across conditions: The pre-critical, critical and spill-over region. Figure \ref{fig:whole_sentence} C shows the pooled reading times of these regions (and additionally the pre-pre-critical region which was identical across conditions as well, but might be confounded by spill-over from the syntactic manipulation). The pre-pre-critical and pre-critical region show longer reading times for high vs.\ low semantic interference conditions. The critical verb was read slowest when syntactic and semantic interference were both high and fastest when syntactic interference was high and semantic interference low. The other two conditions showed intermediate reading times. The spill-over region again primarily showed a difference between high and low semantic interference conditions.

>>>>

\paragraph{Bayes factors for model comparison}

In order to quantify the uncertainty on the parameters of interest, we report 95\% credible intervals.  
These intervals represent the range over which we can be 95\% certain that the values of the parameter lies, given the statistical model and the data. However, formal hypothesis testing cannot be carried out without a likelihood ratio test that compares two alternative models \parencite{schad_etal_2022_BF,Royall}. For this reason, formal hypothesis tests for the presence or absence of the effects of interest were carried out using Bayes factors. Because Bayes factors can be very sensitive to prior specifications on the target parameter being tested \parencite{schad_etal_2022_BF}, 
we report a sensitivity analysis using a range of prior specifications for the relevant parameters (see Table \ref{tab:spr_priors}). The priors assume a priori effect sizes for the main effects of syntactic and semantic interference, and their interaction, ranging from 0-8 ms, 0-40 ms, or 0-81 ms.  These priors are based on the observed effect sizes and their uncertainties in previous reading studies that use the present design \parencite{vandyke07,mertzen}.

@@

Bayes factors can be interpreted as follows \parencite[e.g., ][]{lee2014bayesian}: if BF$_{10}$ > 1, it provides evidence in favor of the effect of interest. If BF$_{10}$ < 1, it provides evidence in favor of the null model. The larger the value of BF$_{10}$, the stronger the evidence for the effect and the smaller the value of BF$_{10}$ is, the stronger the evidence for the null. 
In general, a large number of iterations is needed in the brms package in order to obtain stable estimates of the Bayes factor  \citep{schad_etal_2022_BF}.
For this reason, models were run with four chains and 20,000 iterations in each chain, with the first 2,000 iterations in each chain being discarded as the warm-up phase.  Bayes factors were unstable when varying slopes were included; for this reason, we fit the models only with varying intercepts by participant and by item.

@@

@@

\subsubsection{Comprehension question accuracy (Experiments 1a and 1b combined)}

After exclusion of participants with accuracy below 70\,\%, overall accuracy (including fillers) was good; on average 85.9\,\% (range: 71.9 - 100\,\%). Accuracy for the critical items was 81.1\,\% (range: 20 - 100\,\%). Condition-wise accuracy is shown in Table \ref{tab:spr_acc}.

\begin{table}[]
    \caption{By-condition accuracy in critical trials in the SPR experiment (E1a and E1b combined).}
    \label{tab:spr_acc}
    \centering
    \begin{tabular}{llr}
    \toprule
    syntactic & semantic & accuracy \%\\
    \midrule
        low &  low & 86.0\\
        low &  high & 77.7\\
        high &  low & 85.8\\
        high &  high & 74.9\\
    \bottomrule
    \end{tabular}
\end{table}

Table \ref{tab:spr_acc_mod} shows the results of the generalized mixed model (log-odds scale) analyzing the comprehension accuracy for the critical items (Experiments 1a and 1b combined). The estimates and 95\,\% credible intervals show primarily a reduction in comprehension accuracy for high compared to low semantic interference conditions.

\begin{table}[H]
    \caption{Results in log-odds from the Bayesian generalized model analyzing the comprehension accuracy in the SPR Experiments 1a and 1b.}
    \label{tab:spr_acc_mod}
    \centering
    \begin{tabular}{lrr}
    \toprule
    & Estimate &  95\% CrI  \\
    \midrule
Intercept& 1.66 &   [1.38, 1.93]\\
syntactic& -0.08 &  [-0.16, -0.01]\\
semantic&  -0.61 & [-0.69, -0.54]\\
interaction& -0.10&  [-0.22, 0.02]\\
    \bottomrule
    \end{tabular}
\end{table}

Since the accuracies are not of primary interest, we do not carry out Bayes factors analyses for these.

\subsubsection{Self-paced reading times (Experiments 1a and 1b combined)}
Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence} A and B, separately for high and low syntactic interference because these conditions had partially different sentence structure. 
%\clearpage

@@

It is apparent from Figure \ref{fig:whole_sentence} A and B that, regardless of the syntactic manipulation  at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions. This is probably due to encoding interference \citep{mertzen, jaeger_etal_2015_retrieval_encoding} and persisted in the following regions. Nonetheless, we refrained from analyzing the distractor and the immediately following regions, because they differed between conditions. Instead we focused our analyses on regions which were identical across conditions: The pre-critical, critical and spill-over region. Figure \ref{fig:whole_sentence} C shows the pooled reading times of these regions (and additionally the pre-pre-critical region which was identical across conditions as well, but might be confounded by spill-over from the syntactic manipulation). The pre-pre-critical and pre-critical region show longer reading times for high vs.\ low semantic interference conditions. The critical verb was read slowest when syntactic and semantic interference were both high and fastest when syntactic interference was high and semantic interference low. The other two conditions showed intermediate reading times. The spill-over region again primarily showed a difference between high and low semantic interference conditions.
